---
title: "24343_Zhigansk_66-46N_123-24E"
output: html_notebook
---
#### 
```{r require, include=FALSE}
require(zoo);require(grnn);require(grt);require(lubridate);require(foreach);require(openxlsx); require(imputeTS)
Sys.setenv(R_ZIPCMD = paste0("C:/Users/IVA/Dropbox/Apps", "/bin/zip.exe"))  ## path to zip.exe
# 
```

```{r utils, echo=FALSE}
is.leapyear=function(year){
  return(((year %% 4 == 0) & (year %% 100 != 0)) | (year %% 400 == 0))
}
setZeroNegativeNumber <- function(n){
  if(n < 0) return(0)
  else return(n)
}
as.zero.negative <- function(vec){
  return(sapply(vec, setZeroNegativeNumber))
}
select.year <- function(years, num){
  year <- years[years$Year == num, ]
  return(year)
}
select.year.prec <- function(years, num){
  year <- select.year(years, num)
  year.prec <- year$PRECIP
  return(year.prec)
}
select.year.temp <- function(years, num){
  year <- select.year(years, num)
  year.temp <- year$TMEAN
  return(year.temp)
}
```
####  Reading file with climatic data in the format of the site aisori.
Convert and burn in my form of CVS

My file name format-write station code and name, and then write its coordinates
```{r read_table.aisori, echo=FALSE}
file.cli.path <- "/home/larisa/Dropbox/Apps/na_grnn_year/cli/24343_Zhigansk_66-46N_123-24E"
#file.cli.path <- "C:/Users/IVA/Dropbox/Apps/na_grnn_year/cli/24343_Zhigansk_66-46N_123-24E"
file.name <- "/24343_Zhigansk.csv"; file.name.xlsx <- "/24343_Zhigansk.xlsx"

aisori.cli <- read.csv(paste0(file.cli.path,file.name), header = FALSE, sep = ";", dec = ".")  # read Climate
aisori.cli <- aisori.cli[-c(5, 7, 9, 11, 13, 14)]  # we delete excess columns
aisori.cli <- setNames(aisori.cli, c("Station", "Year", "Month", "Day", "TMIN", "TMEAN", "TMAX", "PRECIP")) # Assign columns names
df.cli.site <- aisori.cli
str(df.cli.site)
summary(df.cli.site)
min.temp <- min(df.cli.site$TMEAN, na.rm=T)
max.temp <- max(df.cli.site$TMEAN, na.rm=T)
min.prec <- min(df.cli.site$PRECIP, na.rm=T)
max.prec <- max(df.cli.site$PRECIP, na.rm=T)

plot(df.cli.site[, c(6)], col="gray")
plot(df.cli.site[, c(8)], col="gray")

```
#### Number of gaps in the data by year

In tabicu output only years with nonzero passes. The second column precipitation third temperature
```{r plot_NA, echo=FALSE}
vec.years <- unique(df.cli.site$Year)
site.prec.na <- 1:length(vec.years); site.temp.na <- 1:length(vec.years); k <- 1
cat("Year, Prec, Temp\n")
for (i in vec.years){
  site.prec.na[k] <- sum(is.na(select.year.prec(df.cli.site, i)))
  site.temp.na[k] <- sum(is.na(select.year.temp(df.cli.site, i)))
  if(site.prec.na[k] != 0 | site.temp.na[k] != 0){
    cat(i, site.prec.na[k], site.temp.na[k], "\n")  
  }
  k <- k + 1
}
require(ggplot2)
plot.prec.na <- data.frame(years=vec.years, prec=site.prec.na)
ggplot(plot.prec.na, aes(x=years, y=prec)) + 
  geom_point(color="steelblue", size=3) + geom_rug() +
  labs(title="plot.prec.na") +
  scale_color_brewer(palette="Paired") + theme_minimal()

plot.temp.na <- data.frame(years=vec.years, temp=site.temp.na)
ggplot(plot.temp.na, aes(x=years, y=temp)) + 
  geom_point(color="#b47846", size=3) + geom_rug() +
  labs(title="plot.temp.na") +
  scale_color_brewer(palette="Paired") + theme_minimal()

```
#### Fill-in temperature and precipitation. There is no data for the entire year.
```{r impute_prec, echo=FALSE}
df.cli.site.period <- df.cli.site[df.cli.site$Year > 1935, ]
summary(df.cli.site.period[, c(6, 8)])

grnn.prec <- function(year.filled, R, K){
  if (is.leapyear(year.filled)) { 
    sum.prec <- rep.int(0, 366)
  } else { 
    sum.prec <- rep.int(0, 365) }
  cat("Length ", year.filled, "year= ", length(sum.prec), "\n")

  for (i in min(df.cli.site.period$Year):max(df.cli.site.period$Year)){
     if (!is.leapyear(i) & is.leapyear(year.filled)){ 
       prec.now <- c(select.year.prec(df.cli.site.period, i), 0.)
     } else if (is.leapyear(i) & !is.leapyear(year.filled)){
        prec.now <- select.year.prec(df.cli.site.period, i)[1:365]
     } else { 
        prec.now <- select.year.prec(df.cli.site.period, i)
     }
     
     for(k in 1:length(sum.prec)){
       if (!is.na(prec.now [k])){
          if (prec.now [k] < 4.) {
              sum.prec[k] <- (sum.prec[k] +  prec.now[k]) / R
          } else if (prec.now [k] > 11.){
             sum.prec[k] <- sum.prec[k] +  prec.now[k] + runif(1, 7, 23)
          } else {
            sum.prec[k] <- (sum.prec[k] +  prec.now[k]) / runif(1, 1, K) 
          }
       }
     }
  }
  for(k in 1:length(sum.prec)){
    if (!is.na(sum.prec [k])){
      if ( sum.prec[k] > max.prec)
        sum.prec[k] <- max.prec
      if (sum.prec[k] < min.prec)
        sum.prec[k] <- min.prec
    } else { stop("*** ERROR grnn.prec - result is.na TRUE") }
  }
  return(sum.prec)
}
grnn.prec.1997 <- grnn.prec(1997, 2, 1.7)
plot(grnn.prec.1997)
grnn.prec.2014 <- grnn.prec(2014, 4, 1.5)
plot(grnn.prec.2014)

```
#### Fill-in temperature and precipitation. There is no data for the entire year.
```{r impute_temp, echo=FALSE}
summary(df.cli.site.period[, c(6, 8)])

grnn.temp <- function(year.filled, R, K){
  if (is.leapyear(year.filled)) { 
    sum.temp <- rep.int(0, 366)
  } else { 
    sum.temp <- rep.int(0, 365) }
  cat("Length ", year.filled, "year= ", length(sum.temp), "\n")

  for (i in min(df.cli.site.period$Year):max(df.cli.site.period$Year)){
     if (!is.leapyear(i) & is.leapyear(year.filled)){ 
       temp.now <- c(select.year.temp(df.cli.site.period, i), 0.)
     } else if (is.leapyear(i) & !is.leapyear(year.filled)){
        temp.now <- select.year.temp(df.cli.site.period, i)[1:365]
     } else {
        temp.now <- select.year.temp(df.cli.site.period, i)
     }
     
     for(k in 1:length(sum.temp)){
       if (!is.na(temp.now [k])){
         sum.temp[k] <- (sum.temp[k] +  temp.now[k]) / runif(1, R, K) 
       }
     }
  }
  for(k in 1:length(sum.temp)){
    if (!is.na(sum.temp [k])){
      if ( sum.temp[k] > max.temp)
        sum.temp[k] <- max.temp
      if (sum.temp[k] < min.temp)
        sum.temp[k] <- min.temp
    } else { stop("*** ERROR grnn.temp - result is.na TRUE") }
  }
  return(sum.temp)
}

grnn.temp.1997 <- grnn.temp(1997, 1.9 ,2.1)
plot(grnn.temp.1997)
grnn.temp.2014 <- grnn.temp(2014, 1.8 ,2.0)
plot(grnn.temp.2014)

```
#### Replacing missing values by year
```{r insert_new_vectors, echo=FALSE}
plot(df.cli.site.period$PRECIP)
df.cli.site.period[df.cli.site.period$Year==1997, 8] <- grnn.prec.1997
df.cli.site.period[df.cli.site.period$Year==2014, 8] <- grnn.prec.2014

df.cli.site.period[df.cli.site.period$Year==1997, 6] <- grnn.temp.1997
df.cli.site.period[df.cli.site.period$Year==2014, 6] <- grnn.temp.2014

plot(df.cli.site.period$PRECIP)
summary(df.cli.site.period$PRECIP)
plot(df.cli.site.period$TMEAN)
summary(df.cli.site.period$TMEAN)
```
#### Application of neural network algorithm for all observations for this station
filled__prec
```{r filled__prec, echo=FALSE}
summary(df.cli.site.period[,c(6,8)])
df.cli.site.prec <- df.cli.site.period$PRECIP
plot(df.cli.site.prec , col="green")
plot(df.cli.site.prec , col="blue")
cli.site.prec.na <- which(is.na(df.cli.site.prec)) # index in vector
cli.site.prec.zoo <- as.zero.negative(zoo::na.spline(df.cli.site.prec))
cli.site.prec.kalman <- as.zero.negative(imputeTS::na.kalman(df.cli.site.prec))
summary(cli.site.prec.zoo)
summary(cli.site.prec.kalman)

for(i in 1:length(cli.site.prec.na)){
  points(cli.site.prec.na[i], cli.site.prec.zoo[cli.site.prec.na[i]], col="red", pch=19)
}
for(i in 1:length(cli.site.prec.na)){
  points(cli.site.prec.na[i], cli.site.prec.kalman[cli.site.prec.na[i]], col="green", pch=19)
}
```
#### Application of neural network algorithm for all observations for this station
filled_temp
```{r filled_temp, echo=FALSE}
cli.site.temp <- df.cli.site.period$TMEAN
plot(cli.site.temp, col="blue")
cli.site.temp.na <- which(is.na(cli.site.temp))
cli.site.temp.zoo <- zoo::na.spline(cli.site.temp)
cli.site.temp.kalman <- imputeTS::na.kalman(cli.site.temp)

summary(cli.site.temp.zoo)
summary(cli.site.temp.kalman)

for(i in 1:length(cli.site.temp.na)){
  points(cli.site.temp.na[i], cli.site.temp.zoo[cli.site.temp.na[i]], col="red", pch=19)
}
for(i in 1:length(cli.site.temp.na)){
  points(cli.site.temp.na[i], cli.site.temp.kalman[cli.site.temp.na[i]], col="green", pch=19)
}
```
#### Write the results in Excel format
```{r create_new_data_cli, echo=FALSE}
require(openxlsx)

df.vso <- data.frame(
  df.cli.site.period[, c(1,2,3,4), ],
  Temp = round(cli.site.temp.kalman, 2),
  Prec = round(cli.site.prec.kalman, 2)
)
summary(df.vso)
openxlsx::write.xlsx(
  df.vso, 
  file = paste0(file.cli.path, file.name.xlsx))
#save(kara_turek, file = paste0(file.cli.path,"/kara_turek.Rdata"))

```

#### Record one year climate data in the format model Vaganova Shashkina
```{r , echo=FALSE}
#Reading climatic data in one year. Format VSO
get_one_year_vso <- function(years, now) {
  one_year <- years[years$Year == now, ]
  cli_dataset <- data.frame(
    one_year[,4], one_year[,3], 
    one_year[,2], as.integer(one_year$Prec*10), 
    as.integer(one_year$Temp*10)
  )
  names(cli_dataset) <- c("day", "month", "year", "prec", "temp");
  return(cli_dataset)
}
#Record one year climate data in the format model Vaganova Shashkina
write.table.vso <- function(year.cli, filled_path, file_name){
  write.table(year.cli, file = paste(filled_path,file_name, sep=""),
              row.names=FALSE, col.names=FALSE, sep="\t")
}

y1939 <- get_one_year_vso(kara_turek, 1939)
summary(y1939)
write.table.vso(y1939, file.cli.path, "/1939.CLI")
y1978 <- get_one_year_vso(kara_turek, 1978)
summary(y1978)
write.table.vso(y1978, file.cli.path, "/1978.CLI")
y1979 <- get_one_year_vso(kara_turek, 1979)
summary(y1979)
write.table.vso(y1979, file.cli.path, "/1979.CLI")
# https://www.rdocumentation.org/packages/BBmisc/versions/1.10
```
