---
title: "36442 kara-turek"
output: html_notebook
---
#### 
  + <https://www.youtube.com/watch?v=JA98rkpT_Dk>
```{r require, include=FALSE}
require(grnn); require(grt); require(openxlsx); require(imputeTS)
Sys.setenv(R_ZIPCMD = paste0("C:/Users/IVA/Dropbox/Apps", "/bin/zip.exe")) 
```

```{r utils}
is.leapyear=function(year){
  return(((year %% 4 == 0) & (year %% 100 != 0)) | (year %% 400 == 0))
}
setZeroNegativeNumber <- function(n){
  if(n < 0) return(0)
  else return(n)
}
as.zero.negative <- function(vec){
  return(sapply(vec, setZeroNegativeNumber))
}
select.year <- function(years, num){
  year <- years[years$Year == num, ]
  return(year)
}
select.year.prec <- function(years, num){
  year <- select.year(years, num)
  year.prec <- year$PRECIP
  return(year.prec)
}
select.year.temp <- function(years, num){
  year <- select.year(years, num)
  year.temp <- year$TMEAN
  return(year.temp)
}
```
```{r na_grnn}
require(grnn); require(grt); require(foreach); require(imputeTS)

na.grnn <- function (x, s) {
  if(!is.null(dim(x)) && dim(x)[2] != 1) {
    stop("Input x is not univariate")
  }
  if (!is.numeric(x)) {
    stop("Input x is not numeric") 
  }
  if (!is.numeric(s)) {
    stop("Input s is not numeric") 
  }
  if (sum(is.na(x)) == 0) return(x)
  ## PRE-PROCESSING DATA
  vec.na <- x; vec.na.scale <- grt::scale(vec.na);  
  vec.na.scale.min <- min(vec.na.scale, na.rm = TRUE); 
  vec.na.scale.max <- max(vec.na.scale, na.rm = TRUE); 
  vec.na.index <- which(is.na(vec.na)) 
  vec.na.scale.na.omit <- na.omit(vec.na.scale);
  days <- 1:length(vec.na); days.scale <- grt::scale(days) 
  days.scale.na.omit <- days.scale[-vec.na.index]; # base::scale(days)
  XY <- data.frame(days.scale.na.omit, vec.na.scale.na.omit) # vec.scale.na.omit
  ##
  L <- grnn::learn(XY, variable.column = ncol(XY))
  grnn <- grnn::smooth(L, sigma = s)
  for (i in vec.na.index) {
    #todo поставить проверку выхода за min max
    G <- grnn::guess(grnn, days.scale[i, 1])
    if (is.na(G)) { G <- 0 }
    vec.na.scale[i] <- G
    #cat("Guess num= ", i, "\n")
  }
  vec.na.unscale <- grt::unscale(vec.na.scale)
  return(as.vector(vec.na.unscale))
} #end na.grnn
na.sample <- function(year.vector, size=36){
  indexs <- sample(length(year.vector), size, replace = FALSE)
  year.vector[indexs] <- NA
  return(year.vector)
}
```
####  Reading file with climatic data in the format of the site aisori. Convert and burn in my form of CVS

My file name format-write station code and name, and then write its coordinates

+ Original format aisori
+ Convert table by extraneous columns and giving the names of the remaining
+ View downloaded data standard
+ Displaying statistics and rendering passes in the data
 
```{r read_table.aisori}
file.cli.path <- "/home/larisa/Dropbox/Apps/na_grnn_year/cli/36442_kara-turek"
#file.cli.path <- "C:/Users/IVA/Dropbox/Apps/na_grnn_year/cli/36442_kara-turek"
file.name <- "/SCH31.txt"; file.name.xlsx <- "/36442_kara-turek.xlsx"

kara_turek.cli <- read.csv(paste0(file.cli.path,file.name), header = FALSE, sep = ";", dec = ".") 
kara_turek.cli <- kara_turek.cli[-c(5, 7, 9, 11, 13, 14)]  # we delete excess columns
kara_turek.cli <- setNames(kara_turek.cli, c("Station", "Year", "Month", "Day", "TMIN", "TMEAN", "TMAX", "PRECIP")) # Assign columns names
df.cli.site <- kara_turek.cli
str(df.cli.site)
summary(df.cli.site)

plotNA.distribution(df.cli.site[, c(6)]); plotNA.distribution(df.cli.site[, c(8)])
plotNA.gapsize(df.cli.site[, c(6)]); plotNA.gapsize(df.cli.site[, c(8)]); 
statsNA(df.cli.site[, c(6)]); cat("\n\n\n"); statsNA(df.cli.site[, c(8)])
min.temp <- min(df.cli.site$TMEAN, na.rm=T)
max.temp <- max(df.cli.site$TMEAN, na.rm=T)
min.prec <- min(df.cli.site$PRECIP, na.rm=T)
max.prec <- max(df.cli.site$PRECIP, na.rm=T)
```
#### Number of gaps in the data by year

In tabicu output only years with nonzero passes. The second column precipitation third temperature
```{r plot_NA, echo=FALSE}
vec.years <- unique(df.cli.site$Year)
site.prec.na <- 1:length(vec.years); site.temp.na <- 1:length(vec.years); k <- 1
for (i in vec.years){
  site.prec.na[k] <- sum(is.na(select.year.prec(df.cli.site, i)))
  site.temp.na[k] <- sum(is.na(select.year.temp(df.cli.site, i)))
  if(site.prec.na[k] != 0 | site.temp.na[k] != 0){
    cat(i, site.prec.na[k], site.temp.na[k], "\n")  
  }
  k <- k + 1
}
require(ggplot2)
plot.prec.na <- data.frame(years=vec.years, prec=site.prec.na)
ggplot(plot.prec.na, aes(x=years, y=prec)) + 
  geom_point(color="steelblue", size=3) + geom_rug() +
  labs(title="plot.prec.na") +
  scale_color_brewer(palette="Paired") + theme_minimal()

plot.temp.na <- data.frame(years=vec.years, temp=site.temp.na)
ggplot(plot.temp.na, aes(x=years, y=temp)) + 
  geom_point(color="#b47846", size=3) + geom_rug() +
  labs(title="plot.temp.na") +
  scale_color_brewer(palette="Paired") + theme_minimal()

```
#### Fill-in temperature and precipitation. There is no data for the entire year.
  + Print-filled year that completely lacked
  + TODO
    + Use the nearby weather station
    + Use tree-ring growth indices
```{r impute_prec, echo=FALSE}
df.cli.site.period <- df.cli.site[df.cli.site$Year > 1938, ]
summary(df.cli.site.period[, c(6, 8)])

these.years <- min(df.cli.site.period$Year):max(df.cli.site.period$Year)

number.days.year <- function (year){
   if (is.leapyear(year)) 
    return(rep.int(0, 366))
  else 
    return(rep.int(0, 365)) 
}
#vector values long as filled year
as.entered <- function (current.year, filled.year, current.year.values){
  if (!is.leapyear(current.year) & is.leapyear(filled.year)){ 
       vec <- c(current.year.values, 0.)
     } else if (is.leapyear(current.year) & !is.leapyear(filled.year)){
        vec <- current.year.values[1:365]
     } else vec <- current.year.values
  return(vec)
}
cut.emissions <- function (generated.vector, min.value, max.value){
  for (k in 1:length(generated.vector)){
    if (!is.na(generated.vector[k])){
      if ( generated.vector[k] > max.value)
        generated.vector[k] <- max.value
      if (generated.vector[k] < min.value)
        generated.vector[k] <- min.value
    } else { stop("*** ERROR cut.emissions - result is.na TRUE") }
  }
  return (generated.vector)
}
grnn.prec <- function(year.filled, R, K){
  sum.prec <- number.days.year(year.filled)
  cat("Length ", year.filled, "year= ", length(sum.prec), "\n")
  for (i in these.years){
     prec.now <- as.entered(i, year.filled, select.year.prec(df.cli.site.period, i))     
     for(k in 1:length(sum.prec)){
       if (!is.na(prec.now [k])){
          if (prec.now [k] < 4.) {
              sum.prec[k] <- (sum.prec[k] +  prec.now[k]) / R
          } else if (prec.now [k] > 11.){
             sum.prec[k] <- sum.prec[k] +  prec.now[k] + runif(1, 7, 23)
          } else {
            sum.prec[k] <- (sum.prec[k] +  prec.now[k]) / runif(1, 1, K) 
          }
       }
     }
  }
  return(cut.emissions(sum.prec, min.prec, max.prec))
}

grnn.prec.1939 <- grnn.prec(1939, 2, 1.7)
plot(grnn.prec.1939, col="red")
grnn.prec.1939[sample(120:250,7)] <- runif(7, 15,35)
points(grnn.prec.1939, col="orange")
grnn.prec.1939.na <- na.sample(grnn.prec.1939)
plot(grnn.prec.1939, col="red")
grnn.prec.1939.grnn <- na.grnn(grnn.prec.1939.na, 0.0025)
points(grnn.prec.1939.grnn, col="green")
points(grnn.prec.1939.na, col="blue")

grnn.prec.1978 <- grnn.prec(1978, 4, 1.5)
plot(grnn.prec.1978)
grnn.prec.1979 <- grnn.prec(1979, 7, 1.9)
plot(grnn.prec.1979)
```
#### Fill-in temperature and precipitation. There is no data for the entire year.
  + Print-filled year that completely lacked
```{r impute_temp, echo=FALSE}
summary(df.cli.site.period[, c(6, 8)])

grnn.temp <- function(year.filled, R, K){
  if (is.leapyear(year.filled)) { 
    sum.temp <- rep.int(0, 366)
  } else { 
    sum.temp <- rep.int(0, 365) }
  cat("Length ", year.filled, "year= ", length(sum.temp), "\n")

  for (i in these.years){
     if (!is.leapyear(i) & is.leapyear(year.filled)){ 
       temp.now <- c(select.year.temp(df.cli.site.period, i), 0.)
     } else if (is.leapyear(i) & !is.leapyear(year.filled)){
        temp.now <- select.year.temp(df.cli.site.period, i)[1:365]
     } else {
        temp.now <- select.year.temp(df.cli.site.period, i)
     }
     
     for(k in 1:length(sum.temp)){
       if (!is.na(temp.now [k])){
         sum.temp[k] <- (sum.temp[k] +  temp.now[k]) / runif(1, R, K) 
       }
     }
  }
  for(k in 1:length(sum.temp)){
    if (!is.na(sum.temp [k])){
      if ( sum.temp[k] > max.temp)
        sum.temp[k] <- max.temp
      if (sum.temp[k] < min.temp)
        sum.temp[k] <- min.temp
    } else { stop("*** ERROR grnn.temp - result is.na TRUE") }
  }
  return(sum.temp)
}

grnn.temp.1939 <- grnn.temp(1939, 1.3 ,2.3)
plot(grnn.temp.1939)
grnn.temp.1978 <- grnn.temp(1978, 1.5 ,2.0)
plot(grnn.temp.1978)
grnn.temp.1979 <- grnn.temp(1979, 1.2 ,2.2)
plot(grnn.temp.1979)

```
#### Replacing missing values by year
+  Before and after. Before and after. Well and distributed so missed by year
```{r insert_new_vectors}
plot(df.cli.site.period$PRECIP)
df.cli.site.period[df.cli.site.period$Year==1939, 8] <- grnn.prec.1939
df.cli.site.period[df.cli.site.period$Year==1978, 8] <- grnn.prec.1978
df.cli.site.period[df.cli.site.period$Year==1979, 8] <- grnn.prec.1979

df.cli.site.period[df.cli.site.period$Year==1939, 6] <- grnn.temp.1939
df.cli.site.period[df.cli.site.period$Year==1978, 6] <- grnn.temp.1978
df.cli.site.period[df.cli.site.period$Year==1979, 6] <- grnn.temp.1979

plot(df.cli.site.period$PRECIP)
plotNA.distribution(df.cli.site.period$PRECIP)
plotNA.distributionBar(df.cli.site.period$PRECIP)
plotNA.gapsize(df.cli.site.period$PRECIP)
print("df.cli.site.period$PRECIP")
summary(df.cli.site.period$PRECIP)
plot(df.cli.site.period$TMEAN)
plotNA.distribution(df.cli.site.period$TMEAN)
plotNA.distributionBar(df.cli.site.period$TMEAN)
plotNA.gapsize(df.cli.site.period$TMEAN)
print("df.cli.site.period$TMEAN")
summary(df.cli.site.period$TMEAN)
```


#### Application of neural network algorithm for all observations for this station 1

```{r filled__prec}
summary(df.cli.site.period[,c(6,8)])
df.cli.site.prec <- df.cli.site.period$PRECIP

plot(df.cli.site.prec , col="blue")
cli.site.prec.na <- which(is.na(df.cli.site.prec)) # index in vector
cli.site.prec.zoo <- as.zero.negative(na.grnn(df.cli.site.prec, 0.001))
cli.site.prec.kalman <- as.zero.negative(
  imputeTS::na.kalman(df.cli.site.prec))
#cli.site.prec.kalman <- as.zero.negative(
#  imputeTS::na.kalman(df.cli.site.prec, model = "auto.arima", smooth = TRUE))

summary(cli.site.prec.zoo)
summary(cli.site.prec.kalman)

for(i in 1:length(cli.site.prec.na)){
  points(cli.site.prec.na[i], cli.site.prec.zoo[cli.site.prec.na[i]], col="red", pch=19)
}
for(i in 1:length(cli.site.prec.na)){
  points(cli.site.prec.na[i], cli.site.prec.kalman[cli.site.prec.na[i]], col="green", pch=19)
}

plotNA.distribution(df.cli.site.prec)
```
#### Application of neural network algorithm for all observations for this station

```{r filled_temp}
cli.site.temp <- df.cli.site.period$TMEAN

plot(cli.site.temp, col="blue")
cli.site.temp.na <- which(is.na(cli.site.temp))
cli.site.temp.zoo <- na.grnn(cli.site.temp, 0.01) 
cli.site.temp.kalman <- imputeTS::na.kalman(cli.site.temp)

summary(cli.site.temp.zoo)
summary(cli.site.temp.kalman)

for(i in 1:length(cli.site.temp.na)){
  points(cli.site.temp.na[i], cli.site.temp.zoo[cli.site.temp.na[i]], col="red", pch=19)
}
for(i in 1:length(cli.site.temp.na)){
  points(cli.site.temp.na[i], cli.site.temp.kalman[cli.site.temp.na[i]], col="green", pch=19)
}

plotNA.distribution(cli.site.temp)
```
#### Write the results in Excel format
```{r create_new_data_cli}
require(openxlsx)

df.vso <- data.frame(
  df.cli.site.period[, c(1,2,3,4), ],
  Temp = round(cli.site.temp.kalman, 2),
  Prec = round(cli.site.prec.kalman, 2)
)
summary(df.vso)
openxlsx::write.xlsx(
  df.vso, 
  file = paste0(file.cli.path, file.name.xlsx))
#save(kara_turek, file = paste0(file.cli.path,"/kara_turek.Rdata"))

```

#### Record one year climate data in the format model Vaganova Shashkina
```{r writeVSO}
#Reading climatic data in one year. Format VSO
get_one_year_vso <- function(years, now) {
  one_year <- years[years$Year == now, ]
  cli_dataset <- data.frame(
    one_year[,4], one_year[,3], 
    one_year[,2], as.integer(one_year$Prec*10), 
    as.integer(one_year$Temp*10)
  )
  names(cli_dataset) <- c("day", "month", "year", "prec", "temp");
  return(cli_dataset)
}
#Record one year climate data in the format model Vaganova Shashkina
write.table.vso <- function(year.cli, filled_path, file_name){
  write.table(year.cli, file = paste(filled_path,file_name, sep=""),
              row.names=FALSE, col.names=FALSE, sep="\t")
}
#
for (i in min(df.cli.site.period$Year):max(df.cli.site.period$Year)){
  cat("Write .CLI file= ", paste0(i,".CLI"), "\n")
  df.vso.i <- get_one_year_vso(df.vso, i)
  summary(df.vso.i)
  write.table.vso(df.vso.i, file.cli.path, paste0("/VSO/", i, ".CLI"))
}
# https://www.rdocumentation.org/packages/BBmisc/versions/1.10
```
