---
title: "filling in missing values in climate data"
output:
  html_document: default
  html_notebook: default
---

```{r read_cli}
rm(list=ls())
#setwd("C:/Users/IVA/Dropbox/24516/trace2017/fmvcd/krest/"); getwd()
setwd("Z:/home/larisa/Dropbox/git/fmvcd/krest/"); getwd()
file_list <- list.files(pattern = "\\.cli$"); file_list
 
for (file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")){
    dataset <- read.table(file, header=FALSE, sep="")
  }
   
  # if the merged dataset does exist, append to it
  if (exists("dataset")){
    temp_dataset <-read.table(file, header=FALSE, sep="")
    dataset<-rbind(dataset, temp_dataset)
    rm(temp_dataset)
  }
}
dataset_krest <- dataset; names(dataset_krest) <- c("day", "month", "year", "prec", "temp")
str(dataset_krest)

```
#### To detect missing data.
```{r filter_cli}
dataset_krest_prec_missing <- dataset_krest[dataset_krest$prec == -9999, c(1,2,3,4)]; length(dataset_krest_prec_missing[,1])
dataset_krest_temp_missing <- dataset_krest[dataset_krest$temp == -9999, c(1,2,3,5)]; length(dataset_krest_temp_missing[,1]); # row.names.data.frame(dataset_krest_temp_missing)
dataset_krest_prec <- dataset_krest[dataset_krest$prec != -9999, c(1,2,3,4)]; summary(dataset_krest_prec)
dataset_krest_temp <- dataset_krest[dataset_krest$temp != -9999, c(1,2,3,5)]; summary(dataset_krest_temp)
ls()

```
#### Replacement of -9999 to a default value NA
```{r set_NA}
dataset_krest[dataset_krest$prec == -9999, 4] <- NA; 
dataset_krest[dataset_krest$temp == -9999, 5] <- NA; summary(dataset_krest)
```
#### replace missing data with appropriate design values.
  + Completely missing at random?
  + Missing at random?
  + Nonrandom pass?
```{r standart_packages}
#install.packages(c("VIM", "mice"))
require(dplyr); require(VIM); require(mice)
```
#### replace missing data with appropriate values calculated using the neural network.
```{r create_train}
#install.packages('neuralnet', dependencies=TRUE, repos='http://cran.rstudio.com/')
library(neuralnet)

train_out <- dataset_krest_temp$temp / 10.0;
train_in <- dataset_krest_temp[, c(1,2,3)]
#train_in <- c(1: length(train_out))
#train_in <- as.numeric(rownames(dataset_krest_temp))
#set.seed(1410) # Make the sample reproducible
#train.sample <- data.frame(NUM=train_in, TEMP=train_out); summary(train.sample)
#train.sample <- dataset_krest_temp
train.sample <- data.frame(train_in, temp=train_out); summary(train.sample)
```
#### Train neuralNet 
```{r train_neuralnet}
rm(net.temp)
train.sample <- train.sample[train.sample$year == 1967,]
net.temp <-  neuralnet(temp ~ day + month + year, train.sample, hidden = c(7,11,7), rep = 1, linear.output = TRUE, threshold = 0.01, stepmax = 10000)

#testdata <- as.numeric(rownames(dataset_krest_temp_missing))
testdata <- dataset_krest_temp_missing[,c(1,2,3)]
testdata <- train.sample[, c(1,2,3)]
net.results <- compute(net.temp, testdata) #Run them through the neural network
ls(net.results)
# Lets see the results
print(head(net.results$net.result))
plot(1:length(train.sample$temp), train.sample$temp)

```
#### syntetic
```{r syntetic}
s_in <- 1:length(train.sample$temp)
s_out <- train.sample$temp
s_out
save(s_out, file='s_out.RData')
tr <- data.frame(X=s_in, Y=s_out)
save(tr, file='tr.RData')
plot(tr)
net.tr<-  neuralnet(formula = 'Y ~ X', data = tr, hidden = c(4,7,4), rep = 1, linear.output = TRUE, threshold = 0.01, stepmax = 100000)
net.res <- compute(net.tr, tr$X)
print(head(net.res$net.result))
plot(net.tr)
```
#### I'm trying to use R's neuralnet package  for prediction.
  + <http://cran.r-project.org/web/packages/neuralnet/neuralnet.pdf>
```{r so}
# http://stats.stackexchange.com/questions/65292/r-neuralnet-compute-give-a-constant-answer
install.packages('neuralnet', dependencies=TRUE, repos='http://cran.rstudio.com/')
library(neuralnet)
Time <- as.integer(runif(50, 0, 100))
Sqrt <- sqrt(Time)
sqrt.data <- data.frame(Time, Sqrt)
plot(sqrt.data)
net.sqrt <- neuralnet(Sqrt~Time, sqrt.data, hidden=10, threshold=0.01)
print(net.sqrt)
plot(net.sqrt)
compute(net.sqrt, (1:10)^2)$net.result

```
#### Prediction - Neural network for regression

Esli vy zanimaetes' modelirovaniem chego ugodno na dannyh, to vam nuzhno imet' dve vyborki: trenirovochnuju i testovuju. Na trenirovochnoj vy stroite model', a na testovoj dolzhny proverit' ee kachestvo.

```{r sin}
install.packages('neuralnet', dependencies=TRUE, repos='http://cran.rstudio.com/')
library(neuralnet)
length(dataset_krest[dataset_krest$year == 1969, 5])
Temp <- dataset_krest[dataset_krest$year == 1969, 5]  / 300.0
Time <- 1:length(Temp)  / 200.0
year1966.data <- data.frame(Time, Temp)
nrow(year1966.data)
nrow(na.omit(year1966.data))
year1966.data.train <- na.omit(year1966.data)
train <- year1966.data.train[sample(nrow(year1966.data.train), 195), ]
plot(train)
net.temp <- neuralnet(Temp~Time, train, hidden=c(7,11,7), threshold=0.001,linear.output=TRUE)
net.temp$result.matrix[1]
plot(net.temp)
# https://journal.r-project.org/archive/2010-1/RJournal_2010-1_Guenther+Fritsch.pdf
# https://www.r-bloggers.com/selecting-the-number-of-neurons-in-the-hidden-layer-of-a-neural-network/
```

#### Links
  + https://www.r-bloggers.com/using-neural-network-for-regression/
  + https://www.r-bloggers.com/fitting-a-neural-network-in-r-neuralnet-package/
  + https://www.r-bloggers.com/a-primer-on-universal-function-approximation-with-deep-learning-in-torch-and-r/
  + https://www.r-bloggers.com/practical-implementation-of-neural-network-based-time-series-stock-prediction-part-1/
  + https://heuristically.wordpress.com/2011/11/17/using-neural-network-for-regression/
  + https://beckmw.wordpress.com/2013/08/12/variable-importance-in-neural-networks/
  + https://beckmw.wordpress.com/tag/neuralnet/

  + http://r-analytics.blogspot.ru/2014/09/r.html#.WHW04dKLRhE
  + http://user2016.org/
  + https://earlconf.com/2017/
  + http://international.sfu-kras.ru/news/17388






