---
title: "filling in missing values in climate data"
output:
  html_document: default
  html_notebook: default
---

```{r read_cli}
rm(list=ls())
#setwd("C:/Users/IVA/Dropbox/24516/trace2017/fmvcd/krest/"); getwd()
#setwd("Z:/home/larisa/Dropbox/git/fmvcd/krest/"); getwd()
setwd("/home/larisa/Dropbox/git/fmvcd/krest/"); getwd()
file_list <- list.files(pattern = "\\.cli$"); file_list

 for (file in file_list){
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")){
    dataset <- read.table(file, header=FALSE, sep="")
  }
   
  # if the merged dataset does exist, append to it
  if (exists("dataset")){
    temp_dataset <-read.table(file, header=FALSE, sep="")
    dataset<-rbind(dataset, temp_dataset)
    rm(temp_dataset)
  }
}
dataset_krest <- dataset; names(dataset_krest) <- c("day", "month", "year", "prec", "temp")
str(dataset_krest)

```
#### Replacement of -9999 to a default value NA
```{r set_NA}
dataset_krest[dataset_krest$prec == -9999, 4] <- NA; 
dataset_krest[dataset_krest$temp == -9999, 5] <- NA; summary(dataset_krest)
```
#### I'm trying to use R's neuralnet package  for prediction.
  + <http://cran.r-project.org/web/packages/neuralnet/neuralnet.pdf>
  + http://gekkoquant.com/2012/05/26/neural-networks-with-r-simple-example/
```{r demo}
# http://stats.stackexchange.com/questions/65292/r-neuralnet-compute-give-a-constant-answer
install.packages('neuralnet', dependencies=TRUE, repos='http://cran.rstudio.com/')
library(neuralnet)
Time <- as.integer(runif(50, 0, 100))
Sqrt <- sqrt(Time)
sqrt.data <- data.frame(Time, Sqrt)
plot(sqrt.data)
net.sqrt <- neuralnet(Sqrt~Time, sqrt.data, hidden=10, threshold=0.01)
print(net.sqrt)
plot(net.sqrt)
compute(net.sqrt, (1:10)^2)$net.result

```
#### Prediction - Neural network for regression

Esli vy zanimaetes' modelirovaniem chego ugodno na dannyh, to vam nuzhno imet' dve vyborki: trenirovochnuju i testovuju. Na trenirovochnoj vy stroite model', a na testovoj dolzhny proverit' ee kachestvo;)

```{r kresr1969}
#install.packages('neuralnet', dependencies=TRUE, repos='http://cran.rstudio.com/')
library(neuralnet)
NORM_TEMP <- 450.0
krestTemp1969 <- dataset_krest[dataset_krest$year == 1969, 5]
summary(krestTemp1969) ###
Temp <- krestTemp1969  / NORM_TEMP
Time <- (1:length(Temp)) / 370.0

year1966.data <- data.frame(Time, Temp)
nrow(year1966.data)
nrow(na.omit(year1966.data))
year1966.data.train <- na.omit(year1966.data)
train <- year1966.data.train[sample(nrow(year1966.data.train), 65), ]
plot(train)
net.temp <- neuralnet(Temp~Time, train, hidden=11, threshold=0.01, linear.output=TRUE)
net.temp$result.matrix[1]
#save(net.temp, file="net.temp65-11.RData")

#plot(net.temp)
# https://journal.r-project.org/archive/2010-1/RJournal_2010-1_Guenther+Fritsch.pdf
# https://www.r-bloggers.com/selecting-the-number-of-neurons-in-the-hidden-layer-of-a-neural-network/
```
####  
  
```{r compute}
computeTemp <- compute(net.temp, train[, c(1)])$net.result
summary(computeTemp)
plot(train[, c(1)], computeTemp, col='red')
points(train, col='blue')

computeTempAll <- compute(net.temp, Time)$net.result
summary(computeTempAll)
#plot(Time, Temp, col='blue')
#points(Time, computeTempAll, col='red')

```
#### filling the gaps and write the result to a file
```{r nnet}
filledKrestTemp1969 <- krestTemp1969
for(i in (1: length(filledKrestTemp1969))){
  if(is.na(filledKrestTemp1969[i])){
    filledKrestTemp1969[i] <- as.integer(computeTempAll[i] * NORM_TEMP)
  }
}
summary(filledKrestTemp1969)
TEMP <- dataset_krest[dataset_krest$year == 1969, 5]
TIME <- 1: length(TEMP)
plot(TIME, TEMP, col='blue')
points(TIME, filledKrestTemp1969, col='red')
filledKrest1969 <- data.frame(dataset_krest[dataset_krest$year == 1969, c(1,2,3,4,5)], filledTemp=filledKrestTemp1969)

```
#### 
```{r save_file}
write.table(filledKrest1969, file="1969.cli.filled", row.names = FALSE)
# write.table(filledKrest1969, file="1969.cli.filled", row.names = FALSE)
```

#### Links
  + https://www.r-bloggers.com/using-neural-network-for-regression/
  + https://www.r-bloggers.com/fitting-a-neural-network-in-r-neuralnet-package/
  + https://www.r-bloggers.com/a-primer-on-universal-function-approximation-with-deep-learning-in-torch-and-r/
  + https://www.r-bloggers.com/practical-implementation-of-neural-network-based-time-series-stock-prediction-part-1/
  + https://heuristically.wordpress.com/2011/11/17/using-neural-network-for-regression/
  + https://beckmw.wordpress.com/2013/08/12/variable-importance-in-neural-networks/
  + https://beckmw.wordpress.com/tag/neuralnet/

  + http://r-analytics.blogspot.ru/2014/09/r.html#.WHW04dKLRhE
  + http://user2016.org/
  + https://earlconf.com/2017/
  + http://international.sfu-kras.ru/news/17388






